---
title: "Single Cell GEX Processing"
fontsize: 11pt
output:
  html_document:
    theme: lumen
    toc: true
    toc_float: 
      collapsed: true
      smooth_scroll: false
    number_sections: false
    highlight: tango
params:
  path_base: "~/Research/compendium"
  data_root: "Data"
  data_id: "SRP372296"
  sample_id: "SRR18922759"
  subject_id: "s01"
  metadata_file: "metaData-Sars-CoV-2.xlsx"
  ig_tr_file: "~/Research/pipeline/Seurat Processing/resources/QC_features_meta_human_v112.rds"
  p10x_file: "SRR18922759/filtered_feature_bc_matrix"
  min.features: 200
  min.cells: 5
  max.mitocondrial.percentage: 20
  is.multi.processed: FALSE
---

# Setup
```{r knitr, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)  # Code and results visible
```

```{r debug-params}
print(params$path_base)
```

## Overview


**Date Created:** June 27, 2024

**Last Updated:** `r Sys.Date()`

**R version:** `r R.Version()$version.string`<br>

**Platform:** `r R.Version()$platform`<br>

**Running under:** `r sessionInfo()$running`

**Notes:**


**To Dos: **
 - Cell annotations using single R <\br>
 
 Ref to these links:  
 
  - https://github.com/dviraran/SingleR
  
  - https://www.singlecellcourse.org/single-cell-rna-seq-analysis-using-seurat.html#basic-quality-control-and-filtering
  
  - https://bioconductor.org/books/release/SingleRBook/using-multiple-references.html
  


## Set the environment

### Path to data and working directory
```{r set environment variables}
#rm(list = ls())

# Set variables for files
path_base <- params$path_base

setwd(path_base)

#path_data_root <- file.path(path_base, params$data_root, params$data_id)
# Removed params$data_id due to differing file structure
path_data_root <- file.path(path_base, params$data_root)
path_to_metadata <- file.path(path_data_root, params$metadata_file)
path_to_10x_file <- file.path(path_data_root, params$p10x_file)
path_to_ig_tr_file <- params$ig_tr_file
#path_to_output_directory <- file.path(path_base, "processed-files", params$data_id)
# Changed "processed-files" to "processed_files"
path_to_output_directory <- file.path(path_base, "processed_files", params$subject_id, params$data_id)
sample_name <- params$sample_id

# Thresholds
min.features = params$min.features
min.cells = params$min.cells
max.mitocondrial.percentage = params$max.mitocondrial.percentage



```


### Load library
```{r}
# list (and install if needed) CRAN packages
# TODO: Add annotation packages
# packages <- c("knitr", "dplyr", "patchwork", "Seurat", "stringr", "tidyverse", "DoubletFinder", "combineSeuratAIRR", "SingleR", "celldex") // removed "combineSeuratAIRR" in this line
packages <- c("knitr", "dplyr", "patchwork", "Seurat", "stringr", "tidyverse", "DoubletFinder", "SingleR", "celldex", "clustree") 
new_pkg <- packages[!(packages %in% installed.packages())]
if (length(new_pkg)) {install.packages(new_pkg)}

# list (and install if needed) Bioconductor packages
packages_additional <- c()
new_pkg_additional <- packages_additional[!(packages_additional %in% installed.packages())]
if (length(new_pkg_additional)) {BiocManager::install(new_pkg_additional)}

# load packages
packages <- sort(append(packages, packages_additional)) # so the list of versions will be in order
for (n in seq_along(packages)) {
  suppressPackageStartupMessages(library(packages[n], character.only = TRUE))
  cat(paste0(packages[n], ": ", packageVersion(packages[n]), "\n")) # print simplified package versions
}
```

### Setting colors and visualization themes

```{r}
color_themes <- c(
"#0CF47A", "#BCF419", "#F4232C", "#CD3E96", "#7B44CD", "#4DCD70", "#A84739", "#277FA9", "#214EA8", "#840071", "#0D0185", "#6F8431", "#665001", "#670D00", "#605669", "#1D1D42", "#462927", "#424124", "#330F00", "#182B32")

mitochondria_scheme <- c("#F55D47", "#7D67FA")
nFeatures_scheme <- c("#47F5A4")
ngenes_scheme <- c("#F5DF47")
```

## Load data

```{r}
# Read metadata
#metadata <- read_metadata(path_to_metadata)
sheet_names <- excel_sheets(path_to_metadata)
metadata <- lapply(sheet_names, function(s) read_excel(path_to_metadata, sheet = s))
names(metadata) <- sheet_names

# Read gene expression
gex_mat <- Read10X(path_to_10x_file)

if (params$is.multi.processed) {
  seurat_obj <- CreateSeuratObject(gex_mat$`Gene Expression`)
} else{
  seurat_obj <- CreateSeuratObject(gex_mat)
}

rm(gex_mat)
```

# Single cell Quality Control

## Filtering cells based on minimum features, expressed in very few cells and mitochondrial gene percentage

QC metrics before filtering
```{r}
seurat_obj[["percent.mt"]] <- PercentageFeatureSet(seurat_obj, pattern = "^MT-")
seurat_obj[["percent.rb"]] <- PercentageFeatureSet(seurat_obj, pattern = "^RP[SL]")
p1 <- VlnPlot(seurat_obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4) 
p1
```





```{r}
p1 <- FeatureScatter(seurat_obj, feature1 = "nCount_RNA", feature2 = "percent.mt") + NoLegend()
p2 <- FeatureScatter(seurat_obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
p1 + p2
```

Filtering step
```{r}
seurat_obj <- subset(seurat_obj, subset = percent.mt < max.mitocondrial.percentage)
seurat_obj <- subset(seurat_obj, subset = nCount_RNA > min.cells)
seurat_obj <- subset(seurat_obj, subset = nFeature_RNA > min.features)
```


QC metrics after filtering
```{r}
seurat_obj[["percent.mt"]] <- PercentageFeatureSet(seurat_obj, pattern = "^MT-")
p1 <- VlnPlot(seurat_obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3) 
p1
```
```{r fig.width=8, fig.height=3}
p1 <- FeatureScatter(seurat_obj, feature1 = "nCount_RNA", feature2 = "percent.mt") + NoLegend()
p2 <- FeatureScatter(seurat_obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") + NoLegend()
p1 + p2
```

## Scaling and Normalization
```{r Normalize data}
# Normalize data
seurat_obj <- NormalizeData(seurat_obj, normalization.method = "LogNormalize", scale.factor = 10000)
```

### Variable feature selection {.tabset}

#### Variable features with IG and TR genes 

```{r variable features}
# Find variable features
seurat_obj <- FindVariableFeatures(seurat_obj, selection.method = "vst", nfeatures = 2000)

top10 <- head(VariableFeatures(seurat_obj), 15)

p1 <- VariableFeaturePlot(seurat_obj) + NoLegend()
p2 <- LabelPoints(plot = p1, points = top10, repel = TRUE)
p1+p2
```

#### Variable features without IG and TR genes 

```{r genes to be excluded}
features_meta <- readRDS(path_to_ig_tr_file)

biotypes_excl <- unique(features_meta[["gene_biotype"]])[grepl(pattern = "^IG_|^TR_",                                                               x = unique(features_meta[["gene_biotype"]]))]
remove_genes <- features_meta[["external_gene_name"]][features_meta[["gene_biotype"]] %in% biotypes_excl]

#remove_genes # List genes to be excluded
```

Now we remove IG and TR features from highly variable features

```{r variable features without ig tr}
# Find variable features
seurat_obj <- FindVariableFeatures(seurat_obj, selection.method = "vst", nfeatures = 2000)

bool_remove_genes <- VariableFeatures(seurat_obj) %in% remove_genes
VariableFeatures(seurat_obj) <- VariableFeatures(seurat_obj)[!bool_remove_genes]

cat("After removing IG/TR genes, the total number of variable features is:", length(VariableFeatures(seurat_obj)),"\n")

top10 <- head(VariableFeatures(seurat_obj), 15)

p1 <- VariableFeaturePlot(seurat_obj) + NoLegend()
p2 <- LabelPoints(plot = p1, points = top10, repel = TRUE, xnudge= 0, ynudge = 0)
p1+p2
```

### Scaling 

```{r}
all.genes <- rownames(seurat_obj)
seurat_obj <- ScaleData(seurat_obj, features= all.genes)
```


## PCA 

```{r}
seurat_obj <- RunPCA(seurat_obj, features = VariableFeatures(object = seurat_obj))
```

### Visualize dimensional loadings

```{r}
# Visualize PCA

print(seurat_obj[["pca"]], dims = 1:5, nfeatures = 5)

VizDimLoadings(seurat_obj, dims =1:2, reduction = "pca")

DimPlot(seurat_obj, reduction = "pca", raster = F)
```

### Heatmap of PCA loadings with respect to genes

```{r eval=FALSE, include=FALSE}
DimHeatmap(seurat_obj , dims = 1, cells = 500, balanced = TRUE)
DimHeatmap(seurat_obj, dims = 1:15, cells = 500, balanced = TRUE)
```

### Elbow Plot

```{r Elbow Plot}
ElbowPlot(seurat_obj, ndims = 60)
```

# Cell Clustering

## Neighbors, Clusters

```{r}
seurat_obj <- FindNeighbors(seurat_obj, dims = 1:50)
suppressMessages(seurat_obj <- FindClusters(seurat_obj, resolution = seq(0.1, 1, 0.1)))
```
## UMAP 

```{r}
seurat_obj <- RunUMAP(seurat_obj, dims = 1:50)
```
```{r}
DimPlot(seurat_obj, reduction = "umap", raster = F, label = TRUE) + NoLegend()
```
## Doublet finder
```{r}
# pK identification (no ground-truth)
sweep.list <- paramSweep(seurat_obj, PCs = 1:50)
sweep.stats <- summarizeSweep(sweep.list)
bcmvn <- find.pK(sweep.stats)
```

```{r}
# Optimal pK is the max of the bomodality coefficent (BCmvn) distribution
bcmvn.max <- bcmvn[which.max(bcmvn$BCmetric),]
optimal.pk <- bcmvn.max$pK
optimal.pk <- as.numeric(levels(optimal.pk))[optimal.pk]
  
## Homotypic doublet proportion estimate
annotations <- seurat_obj@meta.data$seurat_clusters
homotypic.prop <- modelHomotypic(annotations) 
nExp.poi <- round(optimal.pk * nrow(seurat_obj@meta.data)) ## Assuming 7.5% doublet formation rate - tailor for your dataset
nExp.poi.adj <- round(nExp.poi * (1 - homotypic.prop))
  
# run DoubletFinder
seurat_obj <- doubletFinder(seu = seurat_obj, 
                            PCs = 1:50, 
                            pK = optimal.pk,
                            nExp = nExp.poi.adj)
```

```{r}
colnames(seurat_obj@meta.data)[18] <- "doublet_finder"
```



## QC check in UMAP {.tabset}

### Percent Mito 
```{r}
FeaturePlot(seurat_obj, features = c("percent.mt"))
```

### nFeature RNA 
```{r}
FeaturePlot(seurat_obj, features = c("nFeature_RNA"))
```

### nCount RNA 
```{r}
FeaturePlot(seurat_obj, features = c("nCount_RNA"))
```

### Percent Ribo


```{r}
FeaturePlot(seurat_obj, features = c("percent.rb"))
```

### Doublet Finder
```{r}
DimPlot(seurat_obj, reduction = "umap",
        raster = F, group.by = "doublet_finder",
        label = T, label.box = T, repel = T) +
  theme(legend.position = "None") + 
  ggtitle("Doublet Finder")
```

## Picking the right resolution for clustering

> **Note:** Clustree and dot plots are intentionally run **before** doublet removal so resolution choices are evaluated on the full, unfiltered cell population.

### Clustree – resolution stability overview

Clustree visualises how cells move between clusters as resolution increases, helping identify a stable, biologically meaningful resolution.

```{r clustree, fig.width=10, fig.height=12, message=FALSE, warning=FALSE}
if (!requireNamespace("clustree", quietly = TRUE)) {
  install.packages("clustree")
}
library(clustree)

clustree(seurat_obj, prefix = "RNA_snn_res.") +
  theme(legend.position = "right") +
  ggtitle("Clustree – cluster stability across resolutions")
```

---

### PBMC marker gene panel

Dot plots below use a curated set of canonical PBMC markers to aid cluster identity assignment at each resolution. X-axis gene labels are **color-coded by cell type** so groups are immediately readable.

```{r pbmc-markers}
# Canonical PBMC marker genes grouped by cell type
pbmc_markers <- list(
  "B cells"       = c("MS4A1", "CD79A", "CD79B", "CD19"),
  "T cells (CD4)" = c("CD3D", "CD3E", "CD4", "IL7R"),
  "T cells (CD8)" = c("CD8A", "CD8B", "GZMB", "PRF1"),
  "NK cells"      = c("GNLY", "NKG7", "NCAM1", "KLRD1"),
  "Monocytes"     = c("CD14", "LYZ", "S100A8", "S100A9", "FCGR3A"),
  "DC"            = c("FCER1A", "CST3", "CD1C", "CLEC9A"),
  "Plasma cells"  = c("IGHG1", "MZB1", "SDC1", "JCHAIN"),
  "Platelets"     = c("PPBP", "PF4", "GP1BA", "ITGA2B"),
  "pDC"           = c("LILRA4", "IL3RA", "CLEC4C")
)

# Flatten preserving group order
marker_genes_flat <- unique(unlist(pbmc_markers))

# Keep only genes present in the Seurat object
marker_genes_flat <- marker_genes_flat[marker_genes_flat %in% rownames(seurat_obj)]

# ── Color palette: one distinct color per cell type ──────────────────────────
celltype_colors <- c(
  "B cells"       = "#2171B5",
  "T cells (CD4)" = "#238B45",
  "T cells (CD8)" = "#74C476",
  "NK cells"      = "#FD8D3C",
  "Monocytes"     = "#E31A1C",
  "DC"            = "#984EA3",
  "Plasma cells"  = "#A65628",
  "Platelets"     = "#F781BF",
  "pDC"           = "#999999"
)

# Build a named vector: gene → cell-type color (only for genes present)
gene_colors <- setNames(
  unlist(mapply(function(genes, ct) {
    present <- genes[genes %in% rownames(seurat_obj)]
    rep(celltype_colors[ct], length(present))
  }, pbmc_markers, names(pbmc_markers))),
  unlist(mapply(function(genes, ct) {
    genes[genes %in% rownames(seurat_obj)]
  }, pbmc_markers, names(pbmc_markers)))
)
# Align to the final flat gene order used by DotPlot
gene_colors <- gene_colors[marker_genes_flat]

# NOTE: colored x-axis labels are applied directly via `color = gene_colors`
# in element_text(). A matching discrete legend is drawn via cowplot below.

make_dotplot <- function(res_col, res_label) {
  Idents(seurat_obj) <- res_col
  p <- DotPlot(seurat_obj, features = marker_genes_flat, dot.scale = 5) +
    RotatedAxis() +
    scale_color_gradient2(
      low = "steelblue", mid = "white", high = "firebrick", midpoint = 0
    ) +
    ggtitle(paste("Canonical PBMC markers –", res_label)) +
    theme(
      axis.text.x = element_text(
        size = 7, face = "bold",
        color = gene_colors[marker_genes_flat]   # color each label by cell type
      )
    )

  # Build a small color-legend for the cell-type groups using a dummy ggplot
  legend_df <- data.frame(
    CellType = factor(names(celltype_colors), levels = names(celltype_colors)),
    x = 1, y = seq_along(celltype_colors)
  )
  legend_plot <- ggplot(legend_df, aes(x = x, y = y, color = CellType)) +
    geom_point(size = 4) +
    scale_color_manual(values = celltype_colors, name = "Cell type") +
    theme_void() +
    theme(
      legend.title      = element_text(size = 8, face = "bold"),
      legend.text       = element_text(size = 7),
      legend.key.size   = unit(0.5, "cm"),
      legend.position   = "right"
    )
  legend_grob <- cowplot::get_legend(legend_plot)

  # Combine dot plot + color legend side by side
  cowplot::plot_grid(p, legend_grob, ncol = 2, rel_widths = c(1, 0.15))
}

# cowplot is typically available via Seurat; install if missing
if (!requireNamespace("cowplot", quietly = TRUE)) install.packages("cowplot")
library(cowplot)
```

### Resolution UMAPs & Dot Plots {.tabset}

#### Res 0.1
```{r res-0-1-umap}
DimPlot(seurat_obj, 
        reduction="umap", raster=F,
        group.by="RNA_snn_res.0.1",
        label = T, label.box = T, repel=T) +
theme(legend.position = "None") +
ggtitle("Res 0.1")
```

```{r res-0-1-dotplot, fig.width=15, fig.height=5}
make_dotplot("RNA_snn_res.0.1", "Res 0.1")
```

#### Res 0.3
```{r res-0-3-umap}
DimPlot(seurat_obj, 
        reduction="umap", raster=F,
        group.by="RNA_snn_res.0.3",
        label = T, label.box = T, repel=T) +
theme(legend.position = "None") +
ggtitle("Res 0.3")
```

```{r res-0-3-dotplot, fig.width=15, fig.height=5}
make_dotplot("RNA_snn_res.0.3", "Res 0.3")
```

#### Res 0.5
```{r res-0-5-umap}
DimPlot(seurat_obj, 
        reduction="umap", raster=F,
        group.by="RNA_snn_res.0.5",
        label = T, label.box = T, repel=T) +
theme(legend.position = "None") +
ggtitle("Res 0.5")
```

```{r res-0-5-dotplot, fig.width=15, fig.height=5}
make_dotplot("RNA_snn_res.0.5", "Res 0.5")
```

#### Res 0.7
```{r res-0-7-umap}
DimPlot(seurat_obj, 
        reduction="umap", raster=F,
        group.by="RNA_snn_res.0.7",
        label = T, label.box = T, repel=T) +
theme(legend.position = "None") +
ggtitle("Res 0.7")
```

```{r res-0-7-dotplot, fig.width=15, fig.height=5}
make_dotplot("RNA_snn_res.0.7", "Res 0.7")
```

#### Res 0.9
```{r res-0-9-umap}
DimPlot(seurat_obj, 
        reduction="umap", raster=F,
        group.by="RNA_snn_res.0.9",
        label = T, label.box = T, repel=T) +
theme(legend.position = "None") +
ggtitle("Res 0.9")
```

```{r res-0-9-dotplot, fig.width=15, fig.height=5}
make_dotplot("RNA_snn_res.0.9", "Res 0.9")
```

#### Res 1
```{r res-1-umap}
DimPlot(seurat_obj, 
        reduction="umap", raster=F,
        group.by="RNA_snn_res.1",
        label = T, label.box = T, repel=T) +
theme(legend.position = "None") +
ggtitle("Res 1.0")
```

```{r res-1-dotplot, fig.width=15, fig.height=5}
make_dotplot("RNA_snn_res.1", "Res 1.0")
```

Remove doublets
```{r}
seurat_obj <- subset(seurat_obj, doublet_finder == "Singlet")
```

# Cell Annotations

## SingleR


```{r}
sce <- as.SingleCellExperiment(DietSeurat(seurat_obj))
sce
```

```{r}
# Load the reference data
hpca.se <- celldex::HumanPrimaryCellAtlasData()
monaco.se <- celldex::MonacoImmuneData()
blueprint.se <- celldex::BlueprintEncodeData()

# Apply Single R
hpca.main <- SingleR(test = sce, ref = hpca.se, labels = hpca.se$label.main, assay.type.test = 1)
hpca.fine <- SingleR(test = sce, ref = hpca.se, labels = hpca.se$label.fine, assay.type.test = 1)

monaco.main <- SingleR(test = sce, ref = monaco.se, labels = monaco.se$label.main, assay.type.test = 1)
monaco.fine <- SingleR(test = sce, ref = monaco.se, labels = monaco.se$label.fine, assay.type.test = 1)

blueprint.main <- SingleR(test = sce, ref = blueprint.se, labels = blueprint.se$label.main, assay.type.test = 1)
blueprint.fine <- SingleR(test = sce, ref = blueprint.se, labels = blueprint.se$label.fine, assay.type.test = 1)
```

```{r}
# Add cell annotations from single R in seurat object
seurat_obj <- AddMetaData(object = seurat_obj, metadata = hpca.main$labels, col.name = "SingleR_hpca_main")
seurat_obj <- AddMetaData(object = seurat_obj, metadata = hpca.fine$labels, col.name = "SingleR_hpca_fine")

seurat_obj <- AddMetaData(object = seurat_obj, metadata = monaco.main$labels, col.name = "SingleR_monaco_main")
seurat_obj <- AddMetaData(object = seurat_obj, metadata = monaco.fine$labels, col.name = "SingleR_monaco_fine")

seurat_obj <- AddMetaData(object = seurat_obj, metadata = blueprint.main$labels, col.name = "SingleR_blueprint_main")
seurat_obj <- AddMetaData(object = seurat_obj, metadata = blueprint.fine$labels, col.name = "SingleR_blueprint_fine")
```

```{r}
DimPlot(seurat_obj, group.by = "SingleR_hpca_main", raster = F, label = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()

DimPlot(seurat_obj, group.by = "SingleR_monaco_main", raster = F, label = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()

DimPlot(seurat_obj, group.by = "SingleR_blueprint_main", raster = F, label = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
```

```{r}
DimPlot(seurat_obj, group.by = "SingleR_hpca_fine", raster = F, label = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()

DimPlot(seurat_obj, group.by = "SingleR_monaco_fine", raster = F, label = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()

DimPlot(seurat_obj, group.by = "SingleR_blueprint_fine", raster = F, label = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
```

## CellTypist

CellTypist is a Python-based tool for automated cell type classification. We interface with it via the `reticulate` package.

```{r celltypist-setup}
# Install reticulate if needed
if (!requireNamespace("reticulate", quietly = TRUE)) install.packages("reticulate")
library(reticulate)

# Install celltypist into the active Python environment if absent
tryCatch({
  reticulate::py_install("celltypist", pip = TRUE)
}, error = function(e) {
  message("celltypist install skipped (may already be present): ", e$message)
})

celltypist   <- reticulate::import("celltypist")
anndata      <- reticulate::import("anndata")
scipy_sparse <- reticulate::import("scipy.sparse")
```

```{r celltypist-export}
# Export log-normalised counts to an AnnData object (cells × genes)
counts_matrix <- GetAssayData(seurat_obj, assay = "RNA", slot = "data")

counts_csr <- scipy_sparse$csr_matrix(
  list(
    data    = counts_matrix@x,
    indices = counts_matrix@i,
    indptr  = counts_matrix@p
  ),
  shape = c(nrow(counts_matrix), ncol(counts_matrix))
)

adata <- anndata$AnnData(
  X   = counts_csr$T$tocsr(),
  obs = reticulate::r_to_py(data.frame(row.names = colnames(seurat_obj))),
  var = reticulate::r_to_py(data.frame(row.names = rownames(seurat_obj)))
)
cat("AnnData object:", adata$n_obs, "cells ×", adata$n_vars, "genes\n")
```

```{r celltypist-predict}
# Download the Immune_All_Low model (fine-grained PBMC annotations)
# majority_voting smooths predictions within each Seurat cluster
model <- celltypist$models$Model$load(model = "Immune_All_Low.pkl")

predictions <- celltypist$annotate(
  adata,
  model           = model,
  majority_voting = TRUE,
  over_clustering = reticulate::r_to_py(
    as.character(seurat_obj@meta.data$seurat_clusters)
  )
)

cat("CellTypist annotation complete.\n")
print(table(reticulate::py_to_r(predictions$predicted_labels$majority_voting)))
```

```{r celltypist-add-metadata}
# Transfer labels back to the Seurat object
seurat_obj <- AddMetaData(
  seurat_obj,
  metadata = reticulate::py_to_r(predictions$predicted_labels$predicted_labels),
  col.name = "CellTypist_predicted"
)
seurat_obj <- AddMetaData(
  seurat_obj,
  metadata = reticulate::py_to_r(predictions$predicted_labels$majority_voting),
  col.name = "CellTypist_majority_voting"
)
```

```{r celltypist-umap, fig.width=14, fig.height=6}
p_ct1 <- DimPlot(seurat_obj,
                 group.by   = "CellTypist_predicted",
                 reduction  = "umap", raster = FALSE,
                 label = TRUE, label.box = TRUE, repel = TRUE, label.size = 3) +
  NoLegend() + ggtitle("CellTypist – Per-Cell Predictions")

p_ct2 <- DimPlot(seurat_obj,
                 group.by   = "CellTypist_majority_voting",
                 reduction  = "umap", raster = FALSE,
                 label = TRUE, label.box = TRUE, repel = TRUE, label.size = 3) +
  NoLegend() + ggtitle("CellTypist – Majority Voting (per Cluster)")

p_ct1 + p_ct2
```

```{r celltypist-freq-table}
ct_freq <- as.data.frame(table(seurat_obj@meta.data$CellTypist_majority_voting))
colnames(ct_freq) <- c("CellType", "Count")
ct_freq <- ct_freq[order(-ct_freq$Count), ]
knitr::kable(ct_freq, row.names = FALSE,
             caption = "CellTypist majority-voting label frequencies")
```


# Cell Count Summary

```{r cell-count-summary}
# Number of cells from CellRanger (raw, before any filtering)
n_cells_cellranger <- length(colnames(Read10X(path_to_10x_file)))

# Number of cells after full processing (post-QC + doublet removal)
n_cells_processed <- ncol(seurat_obj)

cat("============================================\n")
cat("         Cell Count Summary\n")
cat("============================================\n")
cat(sprintf("Cells from CellRanger output : %d\n", n_cells_cellranger))
cat(sprintf("Cells after QC & processing  : %d\n", n_cells_processed))
cat(sprintf("Cells removed                : %d (%.1f%%)\n",
            n_cells_cellranger - n_cells_processed,
            (n_cells_cellranger - n_cells_processed) / n_cells_cellranger * 100))
cat("============================================\n")
```

# Save Seurat object
```{r}
# Save the seurat object
saveRDS(seurat_obj, file.path(path_to_output_directory, paste0(sample_name, "_processed_object.rds")))

```

# Save B Cells
```{r}
# Subset B cells based on SingleR label (main or fine)
b_cells_hpca_main <- subset(seurat_obj, subset = SingleR_hpca_main == "B_cell")
b_cells_monaco_main <- subset(seurat_obj, subset = SingleR_monaco_main == "B cells")
b_cells_blueprint_main <- subset(seurat_obj, subset = SingleR_blueprint_main == "B-cells")

# Save the subsetted Seurat object
saveRDS(b_cells_hpca_main, file.path(path_to_output_directory, paste0(sample_name, "_Bcells_hpca_main.rds")))
saveRDS(b_cells_monaco_main, file.path(path_to_output_directory, paste0(sample_name, "_Bcells_monaco_main.rds")))
saveRDS(b_cells_blueprint_main, file.path(path_to_output_directory, paste0(sample_name, "_Bcells_blueprint_main.rds")))
```


# Clear up memory
```{r}
rm(list = ls())
gc()
```

