---
title: "Single Cell GEX Processing"
fontsize: 11pt
output:
  html_document:
    theme: lumen
    toc: true
    toc_float: 
      collapsed: true
      smooth_scroll: false
    number_sections: false
    highlight: tango
params:
  path_base: "/mnt/data/Wang"
  data_root: "data"
  data_id: "F09_0"
  sample_id: "F09_0"
  subject_id: "F09"
  metadata_file: "wang_processing_metadata.xlsx"
  ig_tr_file: "/mnt/data/Wang/pipeline/Seurat Pipeline/resources/QC_features_meta_human_v112.rds"
  p10x_file: "F09/F09_0/cell_ranger_F09_0/outs/filtered_feature_bc_matrix"
  min.features: 200
  min.cells: 5
  max.mitocondrial.percentage: 20
  is.multi.processed: FALSE
---

# Setup
```{r knitr, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)  # Code and results visible
# Set the knitr root directory so setwd() inside chunks doesn't trigger a warning
knitr::opts_knit$set(root.dir = normalizePath(params$path_base))
```

```{r debug-params}
print(params$path_base)
```

## Overview


**Date Created:** June 27, 2024

**Last Updated:** `r Sys.Date()`

**R version:** `r R.Version()$version.string`<br>

**Platform:** `r R.Version()$platform`<br>

**Running under:** `r sessionInfo()$running`

**Notes:**


**To Dos: **
 - Cell annotations using single R <\br>
 
 Ref to these links:  
 
  - https://github.com/dviraran/SingleR
  
  - https://www.singlecellcourse.org/single-cell-rna-seq-analysis-using-seurat.html#basic-quality-control-and-filtering
  
  - https://bioconductor.org/books/release/SingleRBook/using-multiple-references.html
  


## Set the environment

### Path to data and working directory
```{r set environment variables}
#rm(list = ls())

# Set variables for files
path_base <- params$path_base

# Note: working directory is set via knitr::opts_knit$set(root.dir) in the setup
# chunk above; setwd() is not used here to avoid knitr warnings.

#path_data_root <- file.path(path_base, params$data_root, params$data_id)
# Removed params$data_id due to differing file structure
path_data_root <- file.path(path_base, params$data_root)
path_to_metadata <- file.path(path_data_root, params$metadata_file)
path_to_10x_file <- file.path(path_data_root, params$p10x_file)
path_to_ig_tr_file <- params$ig_tr_file
#path_to_output_directory <- file.path(path_base, "processed-files", params$data_id)
# Changed "processed-files" to "processed_files"
path_to_output_directory <- file.path(path_base, "processed_files", params$subject_id, params$data_id)
sample_name <- params$sample_id

# Thresholds
min.features = params$min.features
min.cells = params$min.cells
max.mitocondrial.percentage = params$max.mitocondrial.percentage



```


### Load library
```{r}
# list (and install if needed) CRAN packages
# TODO: Add annotation packages
# packages <- c("knitr", "dplyr", "patchwork", "Seurat", "stringr", "tidyverse", "DoubletFinder", "combineSeuratAIRR", "SingleR", "celldex") // removed "combineSeuratAIRR" in this line
packages <- c("knitr", "dplyr", "patchwork", "Seurat", "stringr", "tidyverse", "DoubletFinder", "SingleR", "celldex", "clustree", "readxl", "cowplot", "qs2", "reticulate") 
new_pkg <- packages[!(packages %in% installed.packages())]
if (length(new_pkg)) {install.packages(new_pkg)}

# list (and install if needed) Bioconductor packages
packages_additional <- c()
new_pkg_additional <- packages_additional[!(packages_additional %in% installed.packages())]
if (length(new_pkg_additional)) {BiocManager::install(new_pkg_additional)}

# load packages
packages <- sort(append(packages, packages_additional)) # so the list of versions will be in order
for (n in seq_along(packages)) {
  suppressPackageStartupMessages(library(packages[n], character.only = TRUE))
  cat(paste0(packages[n], ": ", packageVersion(packages[n]), "\n")) # print simplified package versions
}
```

### Setting colors and visualization themes

```{r}
color_themes <- c(
"#0CF47A", "#BCF419", "#F4232C", "#CD3E96", "#7B44CD", "#4DCD70", "#A84739", "#277FA9", "#214EA8", "#840071", "#0D0185", "#6F8431", "#665001", "#670D00", "#605669", "#1D1D42", "#462927", "#424124", "#330F00", "#182B32")

mitochondria_scheme <- c("#F55D47", "#7D67FA")
nFeatures_scheme <- c("#47F5A4")
ngenes_scheme <- c("#F5DF47")
```

## Load data

```{r}
# Read metadata
#metadata <- read_metadata(path_to_metadata)
sheet_names <- excel_sheets(path_to_metadata)
metadata <- lapply(sheet_names, function(s) read_excel(path_to_metadata, sheet = s))
names(metadata) <- sheet_names

# Read gene expression
gex_mat <- Read10X(path_to_10x_file)

if (params$is.multi.processed) {
  seurat_obj <- CreateSeuratObject(gex_mat$`Gene Expression`)
} else{
  seurat_obj <- CreateSeuratObject(gex_mat)
}

rm(gex_mat)
```

# Single cell Quality Control

## Filtering cells based on minimum features, expressed in very few cells and mitochondrial gene percentage

QC metrics before filtering
```{r}
seurat_obj[["percent.mt"]] <- PercentageFeatureSet(seurat_obj, pattern = "^MT-")
seurat_obj[["percent.rb"]] <- PercentageFeatureSet(seurat_obj, pattern = "^RP[SL]")
p1 <- VlnPlot(seurat_obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), ncol = 4) 
p1
```





```{r}
p1 <- FeatureScatter(seurat_obj, feature1 = "nCount_RNA", feature2 = "percent.mt") + NoLegend()
p2 <- FeatureScatter(seurat_obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
p1 + p2
```

Filtering step
```{r}
seurat_obj <- subset(seurat_obj, subset = percent.mt < max.mitocondrial.percentage)
seurat_obj <- subset(seurat_obj, subset = nCount_RNA > min.cells)
seurat_obj <- subset(seurat_obj, subset = nFeature_RNA > min.features)
```


QC metrics after filtering
```{r}
seurat_obj[["percent.mt"]] <- PercentageFeatureSet(seurat_obj, pattern = "^MT-")
p1 <- VlnPlot(seurat_obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3) 
p1
```
```{r fig.width=8, fig.height=3}
p1 <- FeatureScatter(seurat_obj, feature1 = "nCount_RNA", feature2 = "percent.mt") + NoLegend()
p2 <- FeatureScatter(seurat_obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") + NoLegend()
p1 + p2
```

## Scaling and Normalization
```{r Normalize data}
# Normalize data
seurat_obj <- NormalizeData(seurat_obj, normalization.method = "LogNormalize", scale.factor = 10000)
```

### Variable feature selection {.tabset}

#### Variable features with IG and TR genes 

```{r variable features}
# Find variable features
seurat_obj <- FindVariableFeatures(seurat_obj, selection.method = "vst", nfeatures = 2000)

top10 <- head(VariableFeatures(seurat_obj), 20)

p1 <- VariableFeaturePlot(seurat_obj) + NoLegend()
p2 <- LabelPoints(plot = p1, points = top10, repel = TRUE)
p1+p2
```

#### Variable features without IG and TR genes 

```{r genes to be excluded}
features_meta <- readRDS(path_to_ig_tr_file)

biotypes_excl <- unique(features_meta[["gene_biotype"]])[grepl(pattern = "^IG_|^TR_",                                                               
                                                               x = unique(features_meta[["gene_biotype"]]))]
remove_genes <- features_meta[["external_gene_name"]][features_meta[["gene_biotype"]] %in% biotypes_excl]

#remove_genes # List genes to be excluded
```

Now we remove IG and TR features from highly variable features

```{r variable features without ig tr}
# Find variable features
seurat_obj <- FindVariableFeatures(seurat_obj, selection.method = "vst", nfeatures = 2000)

bool_remove_genes <- VariableFeatures(seurat_obj) %in% remove_genes
VariableFeatures(seurat_obj) <- VariableFeatures(seurat_obj)[!bool_remove_genes]

cat("After removing IG/TR genes, the total number of variable features is:", length(VariableFeatures(seurat_obj)),"\n")

top10 <- head(VariableFeatures(seurat_obj), 15)

p1 <- VariableFeaturePlot(seurat_obj) + NoLegend()
p2 <- LabelPoints(plot = p1, points = top10, repel = TRUE, xnudge = 0, ynudge = 0)
p1+p2
```

### Scaling 

```{r}
all.genes <- rownames(seurat_obj)
seurat_obj <- ScaleData(seurat_obj, features= all.genes)
```


## PCA 

```{r}
seurat_obj <- RunPCA(seurat_obj, features = VariableFeatures(object = seurat_obj))
```

### Visualize dimensional loadings

```{r}
# Visualize PCA

print(seurat_obj[["pca"]], dims = 1:5, nfeatures = 5)

VizDimLoadings(seurat_obj, dims =1:2, reduction = "pca")

DimPlot(seurat_obj, reduction = "pca", raster = F)
```

### Heatmap of PCA loadings with respect to genes

```{r eval=FALSE, include=FALSE}
DimHeatmap(seurat_obj , dims = 1, cells = 500, balanced = TRUE)
DimHeatmap(seurat_obj, dims = 1:15, cells = 500, balanced = TRUE)
```

### Elbow Plot

```{r Elbow Plot}
ElbowPlot(seurat_obj, ndims = 60)
```

# Cell Clustering

## Neighbors, Clusters

```{r}
seurat_obj <- FindNeighbors(seurat_obj, dims = 1:50)
suppressMessages(seurat_obj <- FindClusters(seurat_obj, resolution = seq(0.1, 1, 0.1)))
```
## UMAP 

```{r}
seurat_obj <- RunUMAP(seurat_obj, dims = 1:50)
```

```{r}
DimPlot(seurat_obj, reduction = "umap", raster = F, label = TRUE) + NoLegend()
```

## Doublet finder
```{r}
# pK identification (no ground-truth)
sweep.list <- paramSweep(seurat_obj, PCs = 1:50)
sweep.stats <- summarizeSweep(sweep.list)
bcmvn <- find.pK(sweep.stats)
```

```{r}
# Optimal pK is the max of the bomodality coefficent (BCmvn) distribution
bcmvn.max <- bcmvn[which.max(bcmvn$BCmetric),]
optimal.pk <- bcmvn.max$pK
optimal.pk <- as.numeric(levels(optimal.pk))[optimal.pk]
  
## Homotypic doublet proportion estimate
annotations <- seurat_obj@meta.data$seurat_clusters
homotypic.prop <- modelHomotypic(annotations) 
nExp.poi <- round(optimal.pk * nrow(seurat_obj@meta.data)) ## Assuming 7.5% doublet formation rate - tailor for your dataset
nExp.poi.adj <- round(nExp.poi * (1 - homotypic.prop))
  
# run DoubletFinder
seurat_obj <- doubletFinder(seu = seurat_obj, 
                            PCs = 1:50, 
                            pK = optimal.pk,
                            nExp = nExp.poi.adj)
```

```{r}
colnames(seurat_obj@meta.data)[18] <- "doublet_finder"
```



## QC check in UMAP {.tabset}

### Percent Mito 
```{r}
FeaturePlot(seurat_obj, features = c("percent.mt"))
```

### nFeature RNA 
```{r}
FeaturePlot(seurat_obj, features = c("nFeature_RNA"))
```

### nCount RNA 
```{r}
FeaturePlot(seurat_obj, features = c("nCount_RNA"))
```

### Percent Ribo


```{r}
FeaturePlot(seurat_obj, features = c("percent.rb"))
```

### Doublet Finder
```{r}
DimPlot(seurat_obj, reduction = "umap",
        raster = F, group.by = "doublet_finder",
        label = T, label.box = T, repel = T) +
  theme(legend.position = "None") + 
  ggtitle("Doublet Finder")
```

## Picking the right resolution for clustering

> **Note:** Clustree and dot plots are intentionally run **before** doublet removal so resolution choices are evaluated on the full, unfiltered cell population.

### Clustree – resolution stability overview

Clustree visualises how cells move between clusters as resolution increases, helping identify a stable, biologically meaningful resolution.

```{r clustree, fig.height=12, fig.width=10}

clustree(seurat_obj, prefix = "RNA_snn_res.") +
  theme(legend.position = "right") +
  ggtitle("Clustree – cluster stability across resolutions")
```

---

### PBMC marker gene panel

Dot plots below use a curated set of canonical PBMC markers to aid cluster identity assignment at each resolution. X-axis gene labels are **color-coded by cell type** so groups are immediately readable.

Cell specific markers:
 - 




```{r pbmc-markers}
# ─────────────────────────────────────────────────────────────────────────────
# Canonical PBMC marker genes — curated from the literature
#
# B cells
#   MS4A1 (CD20), CD79A, CD79B, CD19 — pan-B identity markers used universally
#   across PBMC scRNA-seq studies. MS4A1/CD20 is absent on plasma cells, making
#   it the canonical gate. Source: Seurat PBMC tutorial (Stuart et al. 2019,
#   Cell); Hao et al. 2021, Cell (Azimuth PBMC reference).
#
# B naive
#   TCL1A — the single most discriminating naive B cell marker in scRNA-seq;
#     absent in memory, plasma cells, and ABCs. Azimuth L2 top marker.
#     Source: Hao et al. 2021, Cell; Stewart et al. 2021, Front. Immunol.
#   IGHD  — encodes IgD heavy chain; transcriptional equivalent of surface IgD
#     used in flow cytometry; robustly detected in 10x 3′ data.
#     Source: Hao et al. 2021; Glass et al. 2020, Cell Reports.
#   IL4R  — consistently naive-enriched in Azimuth and CellTypist references;
#     validated in the INS study defining mature naive B cells.
#     Source: Hao et al. 2021; Sutton et al. 2021, JCI Insight.
#   YBX3  — RNA-binding protein marking naive quiescence.
#     Source: Hao et al. 2021 (Azimuth L2 marker); CellTypist Immune_All_Low.
#
# B memory
#   COCH  — top memory-specific marker in Azimuth; essentially absent in naive
#     and ABC clusters; highly specific to switched memory.
#     Source: Hao et al. 2021, Cell; Stewart et al. 2021, Front. Immunol.
#   AIM2  — enriched in classical memory vs. ABCs and naive; present in Allen
#     Institute Immune Health Atlas and PanglaoDB memory B cell entries.
#     Source: Hao et al. 2021; Allen Inst. Immunology Atlas (2023).
#   TNFRSF13B (TACI) — defines the unswitched/intermediate memory gate in
#     Azimuth ("B intermediate"); used by Glass et al. 2023 to annotate all
#     six memory B cell clusters in tonsillar scRNA-seq.
#     Source: Hao et al. 2021; Glass et al. 2023, JCI Insight.
#   IGHG1 — most abundant IgG subclass transcript; unambiguous class-switched
#     memory marker; also marks plasmablasts (interpret with MS4A1).
#     Source: Glass et al. 2023, JCI Insight; Stewart et al. 2021.
#
# B ABCs (age-associated / atypical memory / DN2)
#   ITGAX (CD11c) — the single best positive ABC marker in scRNA-seq; directly
#     regulated by ZEB2 at the chromatin level; robustly detected in 10x data.
#     Source: Jenks et al. 2018, Immunity; Dai et al. 2024, Science;
#             Woodruff et al. 2020, Nature Immunology.
#   FCRL5 — inhibitory Fc receptor-like protein defining functionally impaired
#     atypical memory B cells across malaria, SLE, and tumor contexts.
#     Source: Kim et al. 2015, PLOS Pathogens; Jenks et al. 2018, Immunity;
#             Pan-cancer B cell atlas 2024, Cancer Cell.
#   ZEB2  — master transcription factor driving ABC lineage commitment; more
#     reliably detected than TBX21 in droplet scRNA-seq; directly represses
#     the GC program and activates ITGAX.
#     Source: Dai et al. 2024, Science; Jenks et al. 2018, Immunity.
#   DUSP4 — one of the most robustly detected ABC markers in scRNA-seq; defines
#     the "B09.DUSP4+ AtM" cluster in the pan-cancer atlas.
#     Source: Pan-cancer B cell atlas 2024, Cancer Cell;
#             RA DN2 B cell study 2023, Front. Immunol.
#   NOTE: TBX21 (T-bet) is the canonical protein-level ABC marker but is
#   notoriously poorly detected in 10x scRNA-seq (mRNA too low); intentionally
#   excluded. Use ITGAX + ZEB2 + DUSP4 as proxy markers.
#
# T cells (CD4)
#   CD3D, CD3E — pan-T cell markers; part of the TCR/CD3 signaling complex.
#     Source: Seurat PBMC tutorial (Stuart et al. 2019); Hao et al. 2021.
#   CD4   — lineage-defining co-receptor for MHC class II.
#     Source: Stuart et al. 2019, Cell; Hao et al. 2021.
#   IL7R  — IL-7 receptor alpha chain; marks naive and memory CD4+ T cells;
#     top differentially expressed gene in CD4 naive cluster in the canonical
#     Seurat PBMC 3k tutorial.
#     Source: Stuart et al. 2019, Cell; Hao et al. 2021.
#
# T cells (CD8)
#   CD8A, CD8B — lineage-defining co-receptors for MHC class I.
#     Source: Stuart et al. 2019; Hao et al. 2021.
#   GZMB  — granzyme B; cytotoxic effector molecule marking activated/effector
#     CD8+ T cells and NK cells.
#     Source: Hao et al. 2021; Szabo et al. 2019, Nature Communications.
#   PRF1  — perforin; pore-forming cytotoxic effector; co-expressed with GZMB
#     in effector CD8 T cells and NK cells.
#     Source: Hao et al. 2021; Stuart et al. 2019 (PBMC 3k DotPlot example).
#
# NK cells
#   GNLY  — granulysin; top NK cell marker in the Seurat PBMC tutorial;
#     also expressed in cytotoxic CD8 T cells.
#     Source: Stuart et al. 2019, Cell; Hao et al. 2021.
#   NKG7  — natural killer cell granule protein 7; robust NK/cytotoxic marker.
#     Source: Stuart et al. 2019; 10x Genomics Loupe Browser PBMC annotation.
#   NCAM1 (CD56) — classical NK cell surface marker; used in 10x Genomics
#     official PBMC annotation guide.
#     Source: 10x Genomics Loupe Browser tutorial; Hao et al. 2021.
#   KLRD1 (CD94) — C-type lectin NK receptor; part of the NKG2 family.
#     Source: Hao et al. 2021 (Azimuth L2 NK marker); PanglaoDB.
#
# Monocytes
#   CD14  — classical monocyte marker (CD14++ monocytes); used in every PBMC
#     scRNA-seq annotation paper as the primary classical monocyte gate.
#     Source: Stuart et al. 2019; Hao et al. 2021; Villani et al. 2017, Science.
#   LYZ   — lysozyme; highly expressed in all monocyte subtypes and DCs;
#     one of the top genes in PC1 in PBMC PCA.
#     Source: Stuart et al. 2019 (PC1 loading); Hao et al. 2021.
#   S100A8, S100A9 — calcium-binding alarmins; top markers of classical
#     monocytes in PBMC scRNA-seq across dozens of studies.
#     Source: Hao et al. 2021; Villani et al. 2017, Science.
#   FCGR3A (CD16) — non-classical monocyte marker; used to distinguish
#     CD14+CD16++ and CD14−CD16++ non-classical monocytes.
#     Source: Stuart et al. 2019; Hao et al. 2021; Villani et al. 2017.
#
# DC (conventional dendritic cells)
#   FCER1A — high-affinity IgE receptor alpha; top cDC2 marker in PBMC scRNA-seq;
#     used in the canonical Seurat tutorial DotPlot.
#     Source: Stuart et al. 2019; Villani et al. 2017, Science.
#   CST3  — cystatin C; expressed in monocytes and DCs; used in the Seurat
#     tutorial as a DC/monocyte shared marker.
#     Source: Stuart et al. 2019; Hao et al. 2021.
#   CD1C  — cDC2 lineage marker (CD1c+ DCs); distinguishes cDC2 from cDC1.
#     Source: Villani et al. 2017, Science; Hao et al. 2021.
#   CLEC9A — cDC1 lineage marker (CLEC9A+ / CD141+ DCs); marks cross-presenting
#     dendritic cells.
#     Source: Villani et al. 2017, Science; Hao et al. 2021.
#
# Plasma cells
#   IGHG1 — IgG1 heavy chain; marks class-switched plasma cells and
#     plasmablasts; highest-expressed Ig in peripheral blood plasma cells.
#     Source: Hao et al. 2021; Glass et al. 2023, JCI Insight.
#   MZB1  — marginal zone B and B1 cell-specific protein; top plasma cell /
#     plasmablast marker in multiple PBMC atlases; used in the Seurat tutorial
#     as a pDC marker (NOTE: subsequent studies clarified MZB1 as plasma cell).
#     Source: Hao et al. 2021; Stuart et al. 2019 (noted in PC12/13).
#   SDC1  (CD138) — syndecan-1; the canonical plasma cell surface marker used
#     in flow cytometry; also detected at mRNA level.
#     Source: Hao et al. 2021; Glass et al. 2023.
#   JCHAIN — immunoglobulin joining chain; marks polymeric IgM/IgA secreting
#     cells; one of the top plasma cell / plasmablast markers in scRNA-seq.
#     Source: Hao et al. 2021; Stuart et al. 2019 (immune_alignment DotPlot).
#
# Platelets (Megakaryocytes)
#   PPBP  — pro-platelet basic protein; top megakaryocyte/platelet marker in
#     PBMC scRNA-seq; used in the Seurat PBMC 3k tutorial.
#     Source: Stuart et al. 2019; Hao et al. 2021.
#   PF4   — platelet factor 4 (CXCL4); co-expressed with PPBP in platelets.
#     Source: Stuart et al. 2019; Hao et al. 2021.
#   GP1BA (CD42b) — GPIb alpha; platelet glycoprotein involved in vWF binding.
#     Source: Hao et al. 2021 (Azimuth platelet markers); PanglaoDB.
#   ITGA2B (CD41) — integrin alpha-IIb; platelet-specific integrin.
#     Source: Hao et al. 2021; PanglaoDB megakaryocyte markers.
#
# pDC (plasmacytoid dendritic cells)
#   LILRA4 — leukocyte immunoglobulin-like receptor A4; top pDC marker in
#     PBMC scRNA-seq; used in Azimuth L2 pDC annotation.
#     Source: Hao et al. 2021; Villani et al. 2017, Science.
#   IL3RA  (CD123) — IL-3 receptor alpha; classical pDC surface marker.
#     Source: Villani et al. 2017, Science; Hao et al. 2021.
#   CLEC4C (CD303/BDCA-2) — C-type lectin receptor; highly specific to pDCs;
#     used as the defining pDC marker across flow cytometry and scRNA-seq.
#     Source: Villani et al. 2017, Science; Hao et al. 2021.
# ─────────────────────────────────────────────────────────────────────────────

pbmc_markers <- list(
  # ── Pan-B identity ──────────────────────────────────────────────────────────
  # Stuart et al. 2019 (Cell); Hao et al. 2021 (Cell / Azimuth PBMC reference)
  "B cells"           = c("MS4A1", "CD79A", "CD79B", "CD19"),

  # ── Naive B cells ───────────────────────────────────────────────────────────
  # TCL1A: Hao et al. 2021; Stewart et al. 2021, Front. Immunol.
  # IGHD:  Glass et al. 2020, Cell Reports; Hao et al. 2021
  # IL4R:  Hao et al. 2021; Sutton et al. 2021, JCI Insight
  # YBX3:  Hao et al. 2021; CellTypist Immune_All_Low model
  "B naive"           = c("TCL1A", "IGHD", "IL4R", "YBX3"),

  # ── Memory B cells ──────────────────────────────────────────────────────────
  # COCH:       Hao et al. 2021; Stewart et al. 2021, Front. Immunol.
  # AIM2:       Hao et al. 2021; Allen Institute Immune Health Atlas 2023
  # TNFRSF13B:  Hao et al. 2021 ("B intermediate"); Glass et al. 2023, JCI Insight
  # IGHG1:      Glass et al. 2023, JCI Insight; Stewart et al. 2021
  "B memory"          = c("COCH", "AIM2", "TNFRSF13B", "IGHG1"),

  # ── Age-associated B cells (ABCs / DN2 / atypical memory) ───────────────────
  # ITGAX:  Jenks et al. 2018, Immunity; Dai et al. 2024, Science
  # FCRL5:  Kim et al. 2015, PLOS Pathogens; Jenks et al. 2018, Immunity
  # ZEB2:   Dai et al. 2024, Science (master TF of ABC lineage)
  # DUSP4:  Pan-cancer B cell atlas 2024, Cancer Cell; RA DN2 study 2023, Front. Immunol.
  # NOTE: TBX21 excluded — poorly detected in 10x scRNA-seq despite being the
  #       canonical protein-level ABC marker (see Cancer Cell 2024 for caveats)
  "B ABCs"            = c("ITGAX", "FCRL5", "ZEB2", "DUSP4"),

  # ── CD4+ T cells ────────────────────────────────────────────────────────────
  # CD3D, CD3E: Stuart et al. 2019 (Cell); Hao et al. 2021
  # CD4:        Stuart et al. 2019; Hao et al. 2021
  # IL7R:       Stuart et al. 2019 (top CD4 naive marker, PBMC 3k tutorial)
  "T cells (CD4)"     = c("CD3D", "CD3E", "CD4", "IL7R"),

  # ── CD8+ T cells ────────────────────────────────────────────────────────────
  # CD8A, CD8B: Stuart et al. 2019; Hao et al. 2021
  # GZMB:       Hao et al. 2021; Szabo et al. 2019, Nature Communications
  # PRF1:       Hao et al. 2021; Stuart et al. 2019 (PBMC 3k DotPlot)
  "T cells (CD8)"     = c("CD8A", "CD8B", "GZMB", "PRF1"),

  # ── NK cells ────────────────────────────────────────────────────────────────
  # GNLY:   Stuart et al. 2019; Hao et al. 2021
  # NKG7:   Stuart et al. 2019; 10x Genomics Loupe Browser PBMC annotation guide
  # NCAM1:  10x Genomics Loupe Browser PBMC guide; Hao et al. 2021
  # KLRD1:  Hao et al. 2021 (Azimuth L2 NK marker)
  "NK cells"          = c("GNLY", "NKG7", "NCAM1", "KLRD1"),

  # ── Monocytes ───────────────────────────────────────────────────────────────
  # CD14:    Stuart et al. 2019; Villani et al. 2017, Science
  # LYZ:     Stuart et al. 2019 (top PC1 loading); Hao et al. 2021
  # S100A8, S100A9: Hao et al. 2021; Villani et al. 2017, Science
  # FCGR3A:  Stuart et al. 2019 (CD16 non-classical gate); Hao et al. 2021
  "Monocytes"         = c("CD14", "LYZ", "S100A8", "S100A9", "FCGR3A"),

  # ── Conventional DCs ────────────────────────────────────────────────────────
  # FCER1A:  Stuart et al. 2019 (canonical DC marker in PBMC 3k DotPlot)
  # CST3:    Stuart et al. 2019; Hao et al. 2021
  # CD1C:    Villani et al. 2017, Science (cDC2 lineage marker)
  # CLEC9A:  Villani et al. 2017, Science (cDC1 / CLEC9A+ DC lineage marker)
  "DC"                = c("FCER1A", "CST3", "CD1C", "CLEC9A"),

  # ── Plasma cells ────────────────────────────────────────────────────────────
  # IGHG1:  Hao et al. 2021; Glass et al. 2023, JCI Insight
  # MZB1:   Hao et al. 2021; Stuart et al. 2019 (noted as plasma/pDC marker)
  # SDC1:   Hao et al. 2021; Glass et al. 2023 (CD138, canonical plasma cell)
  # JCHAIN: Hao et al. 2021; Stuart et al. 2019 (immune_alignment DotPlot)
  "Plasma cells"      = c("XBP1", "MZB1", "SDC1", "JCHAIN"),

  # ── Platelets / Megakaryocytes ───────────────────────────────────────────────
  # PPBP:   Stuart et al. 2019 (top platelet marker, PBMC 3k tutorial)
  # PF4:    Stuart et al. 2019; Hao et al. 2021
  # GP1BA:  Hao et al. 2021 (Azimuth platelet markers); PanglaoDB
  # ITGA2B: Hao et al. 2021; PanglaoDB megakaryocyte markers
  "Platelets"         = c("PPBP", "PF4", "GP1BA", "ITGA2B"),

  # ── Plasmacytoid DCs ────────────────────────────────────────────────────────
  # LILRA4:  Hao et al. 2021; Villani et al. 2017, Science
  # IL3RA:   Villani et al. 2017, Science (CD123, classical pDC gate)
  # CLEC4C:  Villani et al. 2017, Science (CD303/BDCA-2; highly pDC-specific)
  "pDC"               = c("LILRA4", "IL3RA", "CLEC4C")
)
```



```{r pbmc-marker-code}
# Flatten preserving group order
marker_genes_flat <- unique(unlist(pbmc_markers))

# Keep only genes present in the Seurat object
marker_genes_flat <- marker_genes_flat[marker_genes_flat %in% rownames(seurat_obj)]

# ── Color palette: one distinct color per cell type ──────────────────────────
celltype_colors <- c(
  "B cells"       = "#2171B5",
  "B naive"       = "#6BAED6",
  "B memory"    = "#4292C6",
  "B ABCs" = "#08519C",
  "T cells (CD4)" = "#238B45",
  "T cells (CD8)" = "#74C476",
  "NK cells"      = "#FD8D3C",
  "Monocytes"     = "#E31A1C",
  "DC"            = "#984EA3",
  "Plasma cells"  = "#A65628",
  "Platelets"     = "#F781BF",
  "pDC"           = "#999999"
)

# Build a named vector: gene → cell-type color (only for genes present)
gene_colors <- setNames(
  unlist(mapply(function(genes, ct) {
    present <- genes[genes %in% rownames(seurat_obj)]
    rep(celltype_colors[ct], length(present))
  }, pbmc_markers, names(pbmc_markers))),
  unlist(mapply(function(genes, ct) {
    genes[genes %in% rownames(seurat_obj)]
  }, pbmc_markers, names(pbmc_markers)))
)
# Align to the final flat gene order used by DotPlot
gene_colors <- gene_colors[marker_genes_flat]

# NOTE: colored x-axis labels are applied directly via `color = gene_colors`
# in element_text(). A matching discrete legend is drawn via cowplot below.

make_dotplot <- function(res_col, res_label) {
  Idents(seurat_obj) <- res_col

  # Recompute group sizes AFTER filtering to genes actually present in the object
  # This ensures dividers and labels align with what DotPlot renders
  group_sizes_present <- sapply(pbmc_markers, function(genes) {
    sum(genes %in% marker_genes_flat)  # marker_genes_flat is already filtered
  })
  # Drop groups with zero genes (absent entirely from this object)
  group_sizes_present <- group_sizes_present[group_sizes_present > 0]

  divider_positions <- cumsum(group_sizes_present[-length(group_sizes_present)]) + 0.5
  label_positions   <- cumsum(group_sizes_present) - group_sizes_present / 2

  n_clusters <- length(unique(Idents(seurat_obj)))

  p <- DotPlot(seurat_obj, features = marker_genes_flat, dot.scale = 5) +
    geom_vline(xintercept = divider_positions,
               linetype = "dashed", color = "grey50", linewidth = 0.4) +
    annotate("text",
             x        = label_positions,
             y        = Inf,
             label    = names(group_sizes_present),
             color    = "black",
             size     = 2.5,
             fontface = "bold",
             hjust    = 0.5,
             vjust    = 1) +
    RotatedAxis() +
    scale_color_gradient2(
      low = "steelblue", mid = "white", high = "firebrick", midpoint = 0
    ) +
    ggtitle(paste("Canonical PBMC markers –", res_label)) +
    theme(
      axis.text.x = element_text(size = 7, face = "bold",
                                 color = gene_colors[marker_genes_flat]),
      plot.margin = margin(t = 5, r = 5, b = 40, l = 5)
    )

  legend_df <- data.frame(
    CellType = factor(names(celltype_colors), levels = names(celltype_colors)),
    x = 1, y = seq_along(celltype_colors)
  )
  legend_plot <- ggplot(legend_df, aes(x = x, y = y, color = CellType)) +
    geom_point(size = 4) +
    scale_color_manual(values = celltype_colors, name = "Cell type") +
    theme_void() +
    theme(
      legend.title    = element_text(size = 8, face = "bold"),
      legend.text     = element_text(size = 7),
      legend.key.size = unit(0.5, "cm"),
      legend.position = "right"
    )
  legend_grob <- cowplot::get_legend(legend_plot)
  cowplot::plot_grid(p, legend_grob, ncol = 2, rel_widths = c(1, 0.15))
}

```

### Resolution UMAPs & Dot Plots {.tabset}

#### Res 0.1
```{r res-0-1-umap}
DimPlot(seurat_obj, 
        reduction="umap", raster=F,
        group.by="RNA_snn_res.0.1",
        label = T, label.box = T, repel = T) +
theme(legend.position = "None") +
ggtitle("Res 0.1")
```

```{r res-0-1-dotplot, fig.width=15, fig.height=5}
make_dotplot("RNA_snn_res.0.1", "Res 0.1")
```

#### Res 0.3
```{r res-0-3-umap}
DimPlot(seurat_obj, 
        reduction="umap", raster=F,
        group.by="RNA_snn_res.0.3",
        label = T, label.box = T, repel = T) +
theme(legend.position = "None") +
ggtitle("Res 0.3")
```

```{r res-0-3-dotplot, fig.width=15, fig.height=5}
make_dotplot("RNA_snn_res.0.3", "Res 0.3")
```

#### Res 0.5
```{r res-0-5-umap}
DimPlot(seurat_obj, 
        reduction="umap", raster=F,
        group.by="RNA_snn_res.0.5",
        label = T, label.box = T, repel = T) +
theme(legend.position = "None") +
ggtitle("Res 0.5")
```

```{r res-0-5-dotplot, fig.width=15, fig.height=5}
make_dotplot("RNA_snn_res.0.5", "Res 0.5")
```

#### Res 0.7
```{r res-0-7-umap}
DimPlot(seurat_obj, 
        reduction="umap", raster=F,
        group.by="RNA_snn_res.0.7",
        label = T, label.box = T, repel = T) +
theme(legend.position = "None") +
ggtitle("Res 0.7")
```

```{r res-0-7-dotplot, fig.width=15, fig.height=5}
make_dotplot("RNA_snn_res.0.7", "Res 0.7")
```

#### Res 0.9
```{r res-0-9-umap}
DimPlot(seurat_obj, 
        reduction="umap", raster=F,
        group.by="RNA_snn_res.0.9",
        label = T, label.box = T, repel = T) +
theme(legend.position = "None") +
ggtitle("Res 0.9")
```

```{r res-0-9-dotplot, fig.width=15, fig.height=5}
make_dotplot("RNA_snn_res.0.9", "Res 0.9")
```

#### Res 1
```{r res-1-umap}
DimPlot(seurat_obj, 
        reduction="umap", raster=F,
        group.by="RNA_snn_res.1",
        label = T, label.box = T, repel = T) +
theme(legend.position = "None") +
ggtitle("Res 1.0")
```

```{r res-1-dotplot, fig.width=15, fig.height=5}
make_dotplot("RNA_snn_res.1", "Res 1.0")
```

Remove doublets
```{r}
seurat_obj <- subset(seurat_obj, doublet_finder == "Singlet")
```


# Cell Annotations

## SingleR


```{r}
sce <- as.SingleCellExperiment(DietSeurat(seurat_obj))
sce
```

```{r}
# Load the reference data
hpca.se <- celldex::HumanPrimaryCellAtlasData()
monaco.se <- celldex::MonacoImmuneData()
blueprint.se <- celldex::BlueprintEncodeData()

# Apply Single R
hpca.main <- SingleR(test = sce, ref = hpca.se, labels = hpca.se$label.main, assay.type.test = 1)
hpca.fine <- SingleR(test = sce, ref = hpca.se, labels = hpca.se$label.fine, assay.type.test = 1)

monaco.main <- SingleR(test = sce, ref = monaco.se, labels = monaco.se$label.main, assay.type.test = 1)
monaco.fine <- SingleR(test = sce, ref = monaco.se, labels = monaco.se$label.fine, assay.type.test = 1)

blueprint.main <- SingleR(test = sce, ref = blueprint.se, labels = blueprint.se$label.main, assay.type.test = 1)
blueprint.fine <- SingleR(test = sce, ref = blueprint.se, labels = blueprint.se$label.fine, assay.type.test = 1)
```

```{r}
# Add cell annotations from single R in seurat object
seurat_obj <- AddMetaData(object = seurat_obj, metadata = hpca.main$labels, col.name = "SingleR_hpca_main")
seurat_obj <- AddMetaData(object = seurat_obj, metadata = hpca.fine$labels, col.name = "SingleR_hpca_fine")

seurat_obj <- AddMetaData(object = seurat_obj, metadata = monaco.main$labels, col.name = "SingleR_monaco_main")
seurat_obj <- AddMetaData(object = seurat_obj, metadata = monaco.fine$labels, col.name = "SingleR_monaco_fine")

seurat_obj <- AddMetaData(object = seurat_obj, metadata = blueprint.main$labels, col.name = "SingleR_blueprint_main")
seurat_obj <- AddMetaData(object = seurat_obj, metadata = blueprint.fine$labels, col.name = "SingleR_blueprint_fine")
```

```{r}
DimPlot(seurat_obj, group.by = "SingleR_hpca_main", raster = F, label = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()

DimPlot(seurat_obj, group.by = "SingleR_monaco_main", raster = F, label = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()

DimPlot(seurat_obj, group.by = "SingleR_blueprint_main", raster = F, label = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
```

```{r}
DimPlot(seurat_obj, group.by = "SingleR_hpca_fine", raster = F, label = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()

DimPlot(seurat_obj, group.by = "SingleR_monaco_fine", raster = F, label = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()

DimPlot(seurat_obj, group.by = "SingleR_blueprint_fine", raster = F, label = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
```

## CellTypist

CellTypist is a Python-based tool for automated cell type classification. We interface with it via the `reticulate` package.

```{r celltypist-setup}

# celltypist and anndata are pre-installed in the system Python via:
#   python3 -m pip install celltypist anndata --break-system-packages
# Point reticulate at that Python explicitly.
reticulate::use_python(Sys.which("python3"), required = TRUE)

celltypist   <- reticulate::import("celltypist")
anndata      <- reticulate::import("anndata")
scipy_sparse <- reticulate::import("scipy.sparse")

cat("CellTypist", reticulate::py_to_r(celltypist$`__version__`), "ready.\n")
```

```{r celltypist-export}
# Export log-normalised counts to an AnnData object (cells × genes)
counts_matrix <- GetAssayData(seurat_obj, assay = "RNA", layer = "data")
x          <- as.numeric(counts_matrix@x)
i_idx      <- as.integer(counts_matrix@i)
p_idx      <- as.integer(counts_matrix@p)
nrows      <- nrow(counts_matrix)   # genes
ncols      <- ncol(counts_matrix)   # cells
cell_names <- colnames(seurat_obj)
gene_names <- rownames(seurat_obj)

reticulate::py_run_string("
import scipy.sparse as sp
import numpy as np
import anndata as ad
import pandas as pd

# Build genes x cells CSC (matches dgCMatrix storage), then convert to CSR
mat = sp.csc_matrix(
    (np.array(r.x,     dtype=np.float32),
     np.array(r.i_idx, dtype=np.int32),
     np.array(r.p_idx, dtype=np.int32)),
    shape=(int(r.nrows), int(r.ncols))   # genes x cells
).T.tocsr()                               # -> cells x genes

print(f'Matrix shape: {mat.shape}')      # should be cells x genes

adata = ad.AnnData(
    X   = mat,
    obs = pd.DataFrame(index=list(r.cell_names)),
    var = pd.DataFrame(index=list(r.gene_names))
)
print(f'AnnData: {adata.n_obs} cells x {adata.n_vars} genes')
")

adata <- reticulate::py$adata
```

```{r celltypist-predict}
# Download the Immune_All_Low model (fine-grained PBMC annotations)
# majority_voting smooths predictions within each Seurat cluster
model <- celltypist$models$Model$load(model = "Immune_All_Low.pkl")

predictions <- celltypist$annotate(
  adata,
  model           = model,
  majority_voting = TRUE,
  over_clustering = reticulate::r_to_py(
    as.character(seurat_obj@meta.data$seurat_clusters)
  )
)

cat("CellTypist annotation complete.\n")
print(table(reticulate::py_to_r(predictions$predicted_labels$majority_voting)))
```

```{r celltypist-add-metadata}
# Transfer labels back to the Seurat object
seurat_obj <- AddMetaData(
  seurat_obj,
  metadata = reticulate::py_to_r(predictions$predicted_labels$predicted_labels),
  col.name = "CellTypist_predicted"
)
seurat_obj <- AddMetaData(
  seurat_obj,
  metadata = reticulate::py_to_r(predictions$predicted_labels$majority_voting),
  col.name = "CellTypist_majority_voting"
)
```

```{r celltypist-umap, fig.width=14, fig.height=6}
p_ct1 <- DimPlot(seurat_obj,
                 group.by   = "CellTypist_predicted",
                 reduction  = "umap", raster = FALSE,
                 label = TRUE, label.box = TRUE, repel = TRUE, label.size = 3) +
  NoLegend() + ggtitle("CellTypist – Per-Cell Predictions")

p_ct2 <- DimPlot(seurat_obj,
                 group.by   = "CellTypist_majority_voting",
                 reduction  = "umap", raster = FALSE,
                 label = TRUE, label.box = TRUE, repel = TRUE, label.size = 3) +
  NoLegend() + ggtitle("CellTypist – Majority Voting (per Cluster)")

p_ct1 + p_ct2
```

```{r celltypist-freq-table}
ct_freq <- as.data.frame(table(seurat_obj@meta.data$CellTypist_majority_voting))
colnames(ct_freq) <- c("CellType", "Count")
ct_freq <- ct_freq[order(-ct_freq$Count), ]
knitr::kable(ct_freq, row.names = FALSE,
             caption = "CellTypist majority-voting label frequencies")
```


## Azimuth

Azimuth maps cells onto the Hao et al. 2021 multimodal PBMC reference (161k cells, 228 proteins), providing L1, L2, and L3 labels with prediction confidence scores. This is the gold-standard reference-based annotation for PBMC data.

```{r azimuth-setup}
if (!requireNamespace("Azimuth", quietly = TRUE)) {
  remotes::install_github("satijalab/azimuth", ref = "master")
}
library(Azimuth)
```

```{r azimuth-run}
# RunAzimuth maps query cells onto the reference UMAP and transfers labels
seurat_obj <- RunAzimuth(seurat_obj, reference = "pbmcref")

# Azimuth adds: predicted.celltype.l1, predicted.celltype.l2, predicted.celltype.l3
# and corresponding .score columns for prediction confidence
cat("Azimuth annotation complete.\n")
cat("\nL1 label frequencies:\n")
print(table(seurat_obj$predicted.celltype.l1))
cat("\nL2 label frequencies:\n")
print(table(seurat_obj$predicted.celltype.l2))
```

```{r azimuth-umap, fig.width=14, fig.height=5}
p_az1 <- DimPlot(seurat_obj,
                 group.by  = "predicted.celltype.l1",
                 reduction = "umap", raster = FALSE,
                 label = TRUE, label.box = TRUE, repel = TRUE,
                 label.size = 3) +
  NoLegend() + ggtitle("Azimuth – L1 labels")

p_az2 <- DimPlot(seurat_obj,
                 group.by  = "predicted.celltype.l2",
                 reduction = "umap", raster = FALSE,
                 label = TRUE, label.box = TRUE, repel = TRUE,
                 label.size = 3) +
  NoLegend() + ggtitle("Azimuth – L2 labels")

p_az1 + p_az2
```

```{r azimuth-score-qc}
# Visualise prediction confidence — low-score cells are unreliably annotated
VlnPlot(seurat_obj,
        features = "predicted.celltype.l1.score",
        group.by = "predicted.celltype.l1",
        pt.size  = 0) +
  RotatedAxis() +
  ggtitle("Azimuth L1 prediction confidence scores") +
  NoLegend()
```

## scType

scType is a fully marker-based classifier that uses our curated `pbmc_markers` gene sets directly — no external reference is needed. It scores each cluster by the scaled expression of positive markers, making the annotation directly consistent with the dot plots.

```{r sctype-setup}
if (!requireNamespace("HGNChelper", quietly = TRUE)) install.packages("HGNChelper")
library(HGNChelper)

# Load scType functions directly from GitHub (no package install needed)
source("https://raw.githubusercontent.com/IanevskiAleksandr/sc-type/master/R/gene_sets_prepare.R")
source("https://raw.githubusercontent.com/IanevskiAleksandr/sc-type/master/R/sctype_score_.R")
```

```{r sctype-run}
# Build the scType marker input from our curated pbmc_markers list.
# scType expects a data frame with columns: tissueType, cellName, geneSymbolmore1
# (positive markers). We convert our list directly.
sctype_markers <- data.frame(
  tissueType    = "Immune system",
  cellName      = rep(names(pbmc_markers), sapply(pbmc_markers, length)),
  geneSymbolmore1 = unlist(pbmc_markers),
  geneSymbolmore2 = "",   # negative markers — none defined here
  stringsAsFactors = FALSE
)

# Scale the full gene matrix (needed for scType scoring)
seurat_obj <- ScaleData(seurat_obj, features = rownames(seurat_obj), verbose = FALSE)

# Score every cell
es_max <- sctype_score(
  scRNAseqData = LayerData(seurat_obj, assay = "RNA", layer = "scale.data"),
  scaled       = TRUE,
  gs           = setNames(
    lapply(names(pbmc_markers), function(ct) pbmc_markers[[ct]]),
    names(pbmc_markers)
  ),
  gs2          = NULL   # no negative markers
)

# Assign the top-scoring cell type per cluster (using seurat_clusters)
sctype_results <- do.call(rbind, lapply(unique(seurat_obj$seurat_clusters), function(cl) {
  cl_cells  <- which(seurat_obj$seurat_clusters == cl)
  cl_scores <- sort(rowSums(es_max[, cl_cells, drop = FALSE]), decreasing = TRUE)
  data.frame(
    cluster       = cl,
    scType_label  = names(cl_scores)[1],
    scType_score1 = cl_scores[1],
    scType_score2 = cl_scores[2],
    # ratio of top score to runner-up — high ratio = confident call
    scType_ratio  = cl_scores[1] / max(cl_scores[2], 1),
    stringsAsFactors = FALSE
  )
}))

sctype_results <- sctype_results[order(as.numeric(as.character(sctype_results$cluster))), ]
print(sctype_results)

# Map cluster-level labels back to individual cells
cluster_label_map <- setNames(sctype_results$scType_label,
                               as.character(sctype_results$cluster))
labels <- cluster_label_map[as.character(seurat_obj$seurat_clusters)]
names(labels) <- colnames(seurat_obj)
seurat_obj <- AddMetaData(seurat_obj, metadata = labels, col.name = "scType_label")
```

```{r sctype-umap}
DimPlot(seurat_obj,
        group.by  = "scType_label",
        reduction = "umap", raster = FALSE,
        label = TRUE, label.box = TRUE, repel = TRUE, label.size = 3) +
  NoLegend() +
  ggtitle("scType – marker-based cluster annotation")
```


# Cell Count Summary

```{r cell-count-summary}
# Number of cells from CellRanger (raw, before any filtering)
n_cells_cellranger <- length(colnames(Read10X(path_to_10x_file)))

# Number of cells after full processing (post-QC + doublet removal)
n_cells_processed <- ncol(seurat_obj)

cat("============================================\n")
cat("         Cell Count Summary\n")
cat("============================================\n")
cat(sprintf("Cells from CellRanger output : %d\n", n_cells_cellranger))
cat(sprintf("Cells after QC & processing  : %d\n", n_cells_processed))
cat(sprintf("Cells removed                : %d (%.1f%%)\n",
            n_cells_cellranger - n_cells_processed,ee
            (n_cells_cellranger - n_cells_processed) / n_cells_cellranger * 100))
cat("============================================\n")
```

# Save Seurat object
```{r}
# Save the seurat object
saveRDS(seurat_obj, file.path(path_to_output_directory, paste0(sample_name, "_processed_object.rds")))

```

# Subset B cells


```{r}
# Subset B cells based on SingleR label (main or fine)
b_cells_hpca_main <- subset(seurat_obj, subset = SingleR_hpca_main == "B_cell")
b_cells_monaco_main <- subset(seurat_obj, subset = SingleR_monaco_main == "B cells")
b_cells_blueprint_main <- subset(seurat_obj, subset = SingleR_blueprint_main == "B-cells")

# Save the subsetted Seurat object
saveRDS(b_cells_hpca_main, file.path(path_to_output_directory, paste0(sample_name, "_Bcells_hpca_main.rds")))
saveRDS(b_cells_monaco_main, file.path(path_to_output_directory, paste0(sample_name, "_Bcells_monaco_main.rds")))
saveRDS(b_cells_blueprint_main, file.path(path_to_output_directory, paste0(sample_name, "_Bcells_blueprint_main.rds")))
```


# Clear up memory
```{r}
rm(list = ls())
gc()
```

